Generic Properties of D-optimal Designs in Infinite Dimensional Bayesian Linear Inverse Problems

### 2023-03-29 latex/vanishing.tex
Paper almost donegit status !

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
\subsection{Overview}
In this section we analyze D-optimal designs when no model error is
present. Sensor clusterization is discussed in section
\ref{section:clusterization}. We first derive a nonlinear eigenvalue
problem in $\obs$ for D-optimal designs in Section
\ref{subsec:eigenproblem}. In Section \ref{subsec:characterization} we
characterize D-optimal designs as ones that make all eigenvalues equal
in the nonlinear eigenvalue problem. Results are summarized in Theorem
\ref{thm:char}.

\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
condition for D-optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
\end{equation}
with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.

@@

%% Corollary \ref{cor:same ev} is important because it allows us to only think

implies we only need to consider $\obs$ such that  
At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we made no
assumption regarding the ordering of $\{\lambda_i\}$, we can denote
the corresponding non-zero eigenvalues of $\obs^*\obs$ by
$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.

@@

\begin{proof}
  Recall from \eqref{eq:postcov}, that the posterior covariance is
  $\postcov^{-1} = \prcov^{-1} +
  \sigma^{-2}\fwd^*\obs^*\obs\fwd$. Part (1) now follows from Lemma
  \ref{lemma:twice woodbury}. Part (2) follows from Propositin
  \ref{prop:kkt} and Corollary \ref{cor:descending}.
\end{proof}

@@

\begin{proof}
  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
  (3) were proved in Corollary \ref{cor:true target}. Part (4) is a
  consequence Corollariy \ref{cor:eigenvalues}, with the caveat that
  we did not show that finding $\obs$ so that $\obs^*\obs$ has the
  desired eigenvalues is feasible. To this end, we utilize Lemma
  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
  with respect to the first $k$ eigenvectors of $\fwd \prcov
  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
  this concludes the proof.
\end{proof}

\begin{corollary}
  In a D-optimal design, the posterior precision for eigenvector $i$
  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. I.e. the posterior
  precision in $\hil_p$ is constant for $1 \leq i \leq k$, where $k
  =\rank \obs$.
\end{corollary}

@@

\pgfplotstableread{
  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
  4         3.5    0      0       0      0      0           0          0.001
}\clusterization

\pgfplotstableread{
  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
  4         3.5    0      0       0      0      0           0          0.001
}\noclusterization

@@

\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
\subsection{Overview}
In this section we use Theorem \ref{thm:char} to explain how sensor
clusterization may arise in various scenarios. We first consider
sequential designs, and then simultaneous designs.

\subsection{Sequential Design}\label{subsec:clusterization sequential}
In a sequential optimal design problem, each measurement is taken
after results of previous measurements are recorded. In each step
$m=1$, so $k=1$. In this case, Theorem \ref{thm:char} implies each
measurement is taken along the direction of the eigenvector of $\fwd
\prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
decrease fast enough (e.g. exponentially), the

, $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
section \ref{subsec:seq vs sim}, the posterior becomes the prior for
the next step. The eigenvectors of the new posterior are the same as
for the previous one. A new observation will be taken in the direction
of eigenvector with smalles eigenvalue. This leads to:
\begin{align*}
  \begin{split}
    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
    \meas_2 \parallel \ev_2  &\text{ o.w. }
  \end{split}
\end{align*}
In the first case, we get clusterization immediately. In the second,
we consider $\meas_3$. Same reasoning leads to the conclusion that the
eigenvectors of the new prior precision are the same. The eigenvalues,
however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
\sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
direction of eigenvector of smallest precision (note that
$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
\begin{align*}
  \begin{split}
    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
    \sigma^2\lambda^{-1}_3 \\
    \meas_3 \parallel  \ev_3  &\text{ o.w. }
  \end{split}
\end{align*}
Again, the first case results in clusterization. The sequential design
proceeds in this fashion. Every $\meas_i$ is taken in the direction
some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
has the smallest eigenvalue of the previous step's posterior precision
operator. If we do not encounter clusterization after $m$ such steps,
this means that observation $\meas_k$ was taken in the direction
$\ev_k$ for $k=1,\dots,m$. Thus:
\begin{align}\label{eq:failure}
  \begin{split}
    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
  \end{split}
\end{align}

But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
grows at most linearly. In section \ref{subsec:abstract OED} we
assumed $\fwd$ is strongly smoothing, which means eigenvalues of
$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
enough, \eqref{eq:failure} will eventually fail and we will observe
sensor clusterization.

\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
In a simultaneous design scenario, the result is not simple as in the
sequential design case. We now see that a design that exhibits
clusterization is as optimal as one that does not. Corollary
\ref{cor:same ev} tells us that we can realize any design we want
(encoded in $\{\eta_i\}$), given the trace constraint for
$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
we only need to make the posterior flat where we take observations
(recall figure \ref{fig:optimal vs not}). We are left with the freedom
of assigning any part of $\eta_i$ to any observation $\meas_j$. Two
possible assignments for the same problem are illustrated in figure
\ref{fig:clusterization}. The first exhibits clusterization, the
second does not. It is important to note that neither arises as a
result of some numerical optimization scheme. Both designs give the
same design criterion $\tar$, because their posteriors are
identical. By Theorem \ref{thm:char}, both are maximizers of the same
D-optimal design problem. One exhibits clusterization, and one does
not. It is likely that a real-life problem is more restricted than the
model we consider here. Such restriction is the likely cause of
clusterization obseved in, e.g., the toy model of section
\ref{subsec:example}. Exactly how this happens is not clear. At this
point it is the subject of speculation and should be further studied
in the future.

@@

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\noclusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Clusterization (left) and non-clusterization (right) in
    simultaneous D-optimal designs. Posterior precision per
    eigenvector after taking observations $\meas_i$ is plotted. Both
    designs are identical for all practical matters --- their
    posteriors are identical. In the left panel we see how an optimal
    design is achieved with repeated observations. $\meas_2 = \meas_3$
    and $\meas_5 = \meas_6$. This is not necessary though, as can be
    seen in the right panel. Both designs achieve the same levels of
    uncertainty. The left exhibits clusterization. The right does not.}
  \label{fig:clusterization}
\end{figure}

%% \free
%% \begin{proof}
%%   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
%%   \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
%%   $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
%%   otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
%%   orthogonal and will be further restricted later. Then $AA^t = U
%%   SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
%%   eigenvectors by construction. If we can choose $V$ such that $A$
%%   also satisfies the unit norm constraints we are done. These
%%   constraints are, for $j=1,\dots,m$:
%%   \begin{equation}\label{eq:V constraints}
%%    1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
%%   \end{equation}
%%   and we can expect to do this since we assumed $\ttr D = m$.

%%   Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
%%   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
%%   to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
%%   construct such $V$ by sequentially inserting zeros in the diagonal
%%   and not destroying zeros we already introduced, starting from the
%%   last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
%%   let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
%%   the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
%%   rotation $R^{(m)} \in \R^{m \times m}$ by
%%   \begin{equation*}
%%     r^{(m)}_{ab} :=
%%     \begin{cases}
%%       1 & a = b \neq p \text{ or } a = b \neq m \\
%%       \cos \theta & a = b = p  \\
%%      -\sin \theta & a = p, b = m\\
%%       \cos \theta & a = b = m \\
%%       \sin \theta & a = m, b = p \\ 
%%       0 & o.w
%%     \end{cases}
%%   \end{equation*}
%%   Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
%%   $p$ rows and columns. We want to choose $\theta$ such that
%%   \begin{equation}\label{eq:mm}
%%     0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
%%     \theta c_{mp} + \sin^2\theta c_{pp},
%%   \end{equation}
%%   and it suffices to choose $\theta$ such that
%%   \begin{equation*}
%%     c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
%%   \end{equation*}
%%   This quadratic in $\cot\theta$ has a real solution, since
%%   $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
%%   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
%%   $R^{(m-1)}$ that leaves row and column $m$ unchanged and
%%   continue introducing zeros to the diagonal. The assumption $\ttr D =
%%   m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
%%   R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
%% \end{proof}

%% \simdiag
%% \begin{proof}
%%   First, enumerate the eigenvalues of $C + \sum_{j=1}^m
%%   \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
%%   indices of the eigenvectors corresponding to $\xi_i$
%%   \begin{equation*}
%%     S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
%%   \end{equation*}
%%   Define further
%%   \begin{equation*}
%%     A_i := \sum_{k \in S_i} \func_k \func_k^*,
%%   \end{equation*}
%%   which is self-adjoint. Two observations are in order. First,
%%   $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
%%   \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
%%   eigenvalue are orthogonal. For $k \in S_i$
%%   \begin{equation}\label{eq:on vi}
%%     \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
%%   \end{equation}
%%   Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
%%   invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
%%     vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
%%   = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
%%   same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
%%   \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
%% \end{proof}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Characterization of D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
In this section we analyze D-optimal designs when no model error is
present. We first derive a nonlinear eigenvalue problem in $\obs$ for
D-optimal designs, and then we characterize D-optimal designs as ones
that make all eigenvalues equal in the nonlinear eigenvalue
problem. Results are summarized in Theorem \ref{thm:char}.

%% \subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
condition for D-optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
\end{equation}
with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.

@@

Corollary \ref{cor:same ev} implies we only need to consider $\obs$
such that at most $m$ of the eigenvectors of $\obs^*\obs$ have a
positive eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we
made no assumption regarding the ordering of $\{\lambda_i\}$, we can
denote the corresponding non-zero eigenvalues of $\obs^*\obs$ by
$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.

@@

\begin{proof}
  Recall from \eqref{eq:postcov}, that the posterior covariance is
  $\postcov^{-1} = \prcov^{-1} +
  \sigma^{-2}\fwd^*\obs^*\obs\fwd$. Part (1) now follows from Lemma
  \ref{lemma:twice woodbury}. Part (2) follows from Proposition
  \ref{prop:kkt} and Corollary \ref{cor:descending}.
\end{proof}

@@

\begin{proof}
  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
  (3) were proved in Corollary \ref{cor:true target}. Part (4) is a
  consequence of Corollary \ref{cor:eigenvalues}, with the caveat that
  we did not show that finding $\obs$ so that $\obs^*\obs$ has the
  desired eigenvalues is feasible. To this end, we utilize Lemma
  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
  with respect to the first $k$ eigenvectors of $\fwd \prcov
  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
  this concludes the proof.
\end{proof}

\begin{corollary}\label{cor:equal eigs}
  In a D-optimal design, the posterior covariance of the pushforward
  $\fwd_{*} \post$ is constant for eigenvectors $1 \leq i \leq k$
  where $k =\rank \obs$.
\end{corollary}

@@

@@

\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
Corollary \ref{cor:same ev} tells us that we can realize any design we
want (encoded in $\{\eta_i\}$), given the trace constraint
$\tr{\obs^*\obs} = m$ is satisfied. From Theorem \ref{thm:char} we
know we only need to make the posterior flat where we take
observations (recall figure \ref{fig:optimal vs not}). We are left
with the freedom of assigning any part of $\eta_i$ to any observation
$\meas_j$. Two possible assignments for the same problem are
illustrated in figure \ref{fig:clusterization}. The first exhibits
clusterization, the second does not. It is important to note that
neither arises as a result of some numerical optimization scheme. Both
designs give the same design criterion $\tar$, because their
posteriors are identical. By Theorem \ref{thm:char}, both are
maximizers of the same D-optimal design problem. One exhibits
clusterization, and one does not. Of course, a real-life problem is
more restricted than the model we consider here. Such restriction is
the likely cause of clusterization obseved in, e.g., the toy model of
section \ref{subsec:example}. Exactly how this happens is not
clear. At this point it is the subject of speculation and should be
further studied in the future.

A computer implementation of Lemma \ref{lemma:free} (see code in
Supplementary material) generates the design presented in left panel
Figure \ref{fig:clusterization}, which exhibits clusterization, over
the clusterization-free design presented in the right panel of Figure
\ref{fig:clusterization}. Modifying the numerical values of the prior
does not result in a clusterization-free design. Thus, the preference
to clusterization of Lemma \ref{lemma:free} is robust and
clusterization is a generic property of the model presented in this
study.

\pgfplotstableread{
  Label    prior  $o_1o_2$  $o_3$   $o_4$   $o_5$    topper
  1         0.1      0.55       1       0       1      0.001
  2         0.2      1.45       0       1       0      0.001
  3         3.5      0          0       0       0      0.001
}\clusterization

\pgfplotstableread{
  Label    prior  $o_1o_2$  $o_3o_4$   $o_5$     topper
  1         0.1       0.95        0.6        1         0.001
  2         0.2       1.0         1.4        0.        0.001
  3         3.5        0          0          0         0.001
}\noclusterization

@@

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=red!60]    table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=black!60]  table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=orange!60] table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
      %% \addplot [fill=cyan!60]   table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
      %% \addplot [fill=purple!60] table [y=$o_7$, meta=Label, x expr=\coordindex] {\clusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      %% \addlegendentry{$o_6$}
      %% \addlegendentry{$o_7$}   
    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\noclusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=red!60]    table [y=$o_3o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=black!60]  table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      %% \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1o_2$}
      \addlegendentry{$o_3o_4$}
      \addlegendentry{$o_5$}
      %% \addlegendentry{$o_4$}
      %% \addlegendentry{$o_5$}
      %% \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Clusterization (left) and non-clusterization (right) in
    D-optimal designs. Posterior precision per eigenvector after
    taking observations $\meas_i$ is plotted. Both designs are
    identical for all practical matters --- their posteriors are
    equal. In the left panel we see how an optimal design is achieved
    with repeated observations. $\meas_3 = \meas_5$. This is not
    necessary though, as can be seen in the right panel. Both designs
    achieve the same levels of uncertainty. The left exhibits
    clusterization. The right does not.}
  \label{fig:clusterization}
\end{figure}

%% may arise in various scenarios. We first consider
%% sequential designs, and then simultaneous designs.

%% \subsection{Sequential Design}\label{subsec:clusterization sequential}
%% In a sequential optimal design problem, a decision is made on each
%% measurement with all previous measurements fixed. In each step $m=1$,
%% so $k=1$. In this case, Theorem \ref{thm:char} implies each
%% measurement is taken along the direction of the eigenvector of $\fwd
%% \prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
%% decrease fast enough (e.g. exponentially), then $\lambda_i^{-1}$
%% increase exponentially. Then, at some point, a measurement will be
%% taken along an already measured eigenvector. This situation is
%% measurement clusterization.

%% , $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
%% section \ref{subsec:seq vs sim}, the posterior becomes the prior for
%% the next step. The eigenvectors of the new posterior are the same as
%% for the previous one. A new observation will be taken in the direction
%% of eigenvector with smalles eigenvalue. This leads to:
%% \begin{align*}
%%   \begin{split}
%%     \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
%%     \meas_2 \parallel \ev_2  &\text{ o.w. }
%%   \end{split}
%% \end{align*}
%% In the first case, we get clusterization immediately. In the second,
%% we consider $\meas_3$. Same reasoning leads to the conclusion that the
%% eigenvectors of the new prior precision are the same. The eigenvalues,
%% however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
%% \sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
%% direction of eigenvector of smallest precision (note that
%% $\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
%% \begin{align*}
%%   \begin{split}
%%     \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
%%     \sigma^2\lambda^{-1}_3 \\
%%     \meas_3 \parallel  \ev_3  &\text{ o.w. }
%%   \end{split}
%% \end{align*}
%% Again, the first case results in clusterization. The sequential design
%% proceeds in this fashion. Every $\meas_i$ is taken in the direction
%% some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
%% has the smallest eigenvalue of the previous step's posterior precision
%% operator. If we do not encounter clusterization after $m$ such steps,
%% this means that observation $\meas_k$ was taken in the direction
%% $\ev_k$ for $k=1,\dots,m$. Thus:
%% \begin{align}\label{eq:failure}
%%   \begin{split}
%%     \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
%%     \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
%%   \end{split}
%% \end{align}

%% But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
%% grows at most linearly. In section \ref{subsec:abstract OED} we
%% assumed $\fwd$ is strongly smoothing, which means eigenvalues of
%% $(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
%% enough, \eqref{eq:failure} will eventually fail and we will observe
%% sensor clusterization.

%% \subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
%% If the eigenvalues $\lambda_i$ decrease quickly enough, we will
%% observe clusterizaion.
